Large-Scale Content-Based Matching of MIDI and Audio Files

### 2015-05-02 paper.tex
Fix spelling

Motivated by these issues, we propose a system which learns a common representation for the audio and MIDI CQTs in a Hamming space.
By replacing constant-Q spectra with bitvectors, we replace the expensive distance computation with an exclusive-or operation followed by simple table lookup.
This is thanks to the fact that computing the exlcusive-or of two bitvectors $a$ and $b$ will yield a bitvector with $1$s in the dimensions where $a$ and $b$ differ and $0$s elsewhere, and that the number of $1$s in all bitvectors of length $D$ can be precomputed and stored in a table of size $2^D$.
In the course of computing our Hamming space representation, we also implicitly downsample the sequences over time, which provides speedups for both similarity matrix and DTW calculation.
Our approach has the additional potential benefit of learning a representation for comparing audio and MIDI constant-Q spectra, rather than simply using the cosine distance.

@@

The hashing networks for each modality were chosen to have the same architecture: A series of alternating convolutional and pooling layers followed by a series of fully-connected layers.
All convolutional and fully-connected layers used rectifier nonlinearities except the final output layer, which (as in \cite{masci2014multimodal}) used a hyperbolic tangent nonlinearity.
This choice allows us to obtain binary hash vectors by testing whether each output unit is greater or less than zero.
The first convolutional layer had 16 filters with a filter size of five beats by twelve semitones, which gives our network some temporal context and octave invariance.
As advocated by \cite{simonyan2014very}, all subsequent convolutional layers had $2^{n + 3}$ 3x3 filters, where $n$ is the depth of the layer.
All pooling layers performed max-pooling, with a pooling size of 2x2.
Finally, as advocated by \cite{he2015delving}, we initialized all weights with normally distributed random variables with zero mean and a standard deviation of $\sqrt{2/n_{in}}$, where $n_{in}$ is the number of inputs to each layer.
Our model was implemented using Theano \cite{bastien2012theano} and \texttt{lasagne}.\footnote{\texttt{https://github.com/Lasagne/Lasagne}}

To ensure good performance, we optimized all model hyperparameters using Whetlab,\footnote{\texttt{http://www.whetlab.com}} a web API which implements black-box Bayesian optimization \cite{snoek2012practical}.
We used Whetlab to optimize the number of convolutional/pooling layers, the number and size of the fully-connected layers, the RMSProp learning rate and decay parameters, and the $\alpha$ and $m$ regularization parameters of $\mathcal{L}$.
As a hyperparameter optimization objective, we used the Bhattacharyya distance approach described above.
The best performing network found by Whetlab had 2 convolutional layers, 2 ``hidden'' fully-connected layers of with 2048 units in addition to a fully-connected output layer, a learning rate of .001 with a decay parameter of .65, and $\alpha = .5$ and $m = 4$.
This hyperparameter configuration yielded the output hash distance distributions for $\mathcal{P}$ and $\mathcal{N}$ shown in Figure \ref{fig:distances}, which resulted in a Bhattacharryya distance of $0.488$.

>>>>

Motivated by these issues, we propose a system which learns a common representation for the audio and MIDI CQTs in a Hamming space.
By replacing constant-Q spectra with bitvectors, we replace the expensive distance computation with an exclusive-or operation followed by simple table lookup.
This is thanks to the fact that computing the exclusive-or of two bitvectors $a$ and $b$ will yield a bitvector with $1$s in the dimensions where $a$ and $b$ differ and $0$s elsewhere, and that the number of $1$s in all bitvectors of length $D$ can be precomputed and stored in a table of size $2^D$.
In the course of computing our Hamming space representation, we also implicitly downsample the sequences over time, which provides speedups for both similarity matrix and DTW calculation.
Our approach has the additional potential benefit of learning a representation for comparing audio and MIDI constant-Q spectra, rather than simply using the cosine distance.

@@

The hashing networks for each modality were chosen to have the same architecture: A series of alternating convolutional and pooling layers followed by a series of fully-connected layers.
All convolutional and fully-connected layers used rectifier nonlinearities except the final output layer, which (as in \cite{masci2014multimodal}) used a hyperbolic tangent nonlinearity.
This choice allows us to obtain binary hash vectors by testing whether each output unit is greater or less than zero.
The first convolutional layer had 16 filters with a filter size of five beats by twelve semitones, which gives our network some temporal context and octave invariance.
As advocated by \cite{simonyan2014very}, all subsequent convolutional layers had $2^{n + 3}$ 3x3 filters, where $n$ is the depth of the layer.
All pooling layers performed max-pooling, with a pooling size of 2x2.
Finally, as advocated by \cite{he2015delving}, we initialized all weights with normally distributed random variables with zero mean and a standard deviation of $\sqrt{2/n_{in}}$, where $n_{in}$ is the number of inputs to each layer.
Our model was implemented using \texttt{theano} \cite{bastien2012theano} and \texttt{lasagne}.\footnote{\texttt{https://github.com/Lasagne/Lasagne}}

To ensure good performance, we optimized all model hyperparameters using Whetlab,\footnote{\texttt{http://www.whetlab.com}} a web API which implements black-box Bayesian optimization \cite{snoek2012practical}.
We used Whetlab to optimize the number of convolutional/pooling layers, the number and size of the fully-connected layers, the RMSProp learning rate and decay parameters, and the $\alpha$ and $m$ regularization parameters of $\mathcal{L}$.
As a hyperparameter optimization objective, we used the Bhattacharyya distance approach described above.
The best performing network found by Whetlab had 2 convolutional layers, 2 ``hidden'' fully-connected layers of with 2048 units in addition to a fully-connected output layer, a learning rate of .001 with a decay parameter of .65, and $\alpha = .5$ and $m = 4$.
This hyperparameter configuration yielded the output hash distance distributions for $\mathcal{P}$ and $\mathcal{N}$ shown in Figure \ref{fig:distances}, which resulted in a Bhattacharyya distance of $0.488$.
